---
title: "Autism Service Compliance"
output: 
  html_document:
    code_folding: hide
params:
   cmh: "Genesee Health System"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

### This report was generated for `r paste0(params$cmh)` on `r Sys.Date()`

```{r}
# Install packages if you need them

# Load required packages
library(tidyverse); library(stringr); library(magrittr); library(lubridate); library(readxl)
library(DT); library(forcats); library(plotly); library(sparkline);library(feather)
library(data.table); library(reshape2); library(dplyr)
```

# Reading in Data

The data for these analyses is drawn from two primary sources: 

- *Service data* from the region's MIX EMR system and 
- *Waiver enrollment* reports generated from the MDHHS WSA system.

First, we read in the services data and process it for analysis.  Note that we are currently reading from .csv file output for this draft analysis.  A repeatable, automated process will rely on connection to database tables updated via an extract-transfer-load (ETL) process.

```{r}

# Read in .csv files as dataframes
csv_path <- "C:/Users/joshh/OneDrive - TBD Solutions LLC/files/Region10/Autism/"
svs <- read_feather(paste0(csv_path,"svs.feather"))

# Filter df by defined parameter (i.e.. CMH name)
svs %<>% filter(PROVIDER_NAME == params$cmh) %>% droplevels()
```

Next, we read in the waiver enrollment data and process it:

```{r}
# Read in .csv files as dataframes
wsa <- read_feather(paste0(csv_path,"wsa.feather"))
wsa %<>% filter(PIHP_CMH_Name == params$cmh) %>% droplevels()
# If re-opened, client is given new 'Case ID' but Medicaid ID (i.e. 'ID') should be same

# Read in data related to authorized units
wsa_ipos <- read_feather(paste0(csv_path,"wsa_ipos.feather"))
wsa_ipos %<>% filter(PIHP_CHM == params$cmh) %>% droplevels()
```

*Please note that the data in this report may not reflect the services of all enrollees that participated in the Autism Waiver Program.  Specifically, enrollee services may not show in the data for more recent periods where claims were not yet submitted.*

---

# Eligible Autism Waiver Enrollees without a Plan of Service

As an initial summary, we look at the percent of persons who are enrolled in the autism waiver (i.e. `Status == "Open"`), but who do not have an Individualized Plan of Service (IPOS).

```{r}
wsa_tbl <-
wsa %>%
  filter(Status == "Open") %>%
  group_by(PIHP_CMH_Name) %>%
  summarize(
    without_IPOS = n() - sum(IPOSExists, na.rm = T),
    all = n()
  ) %>%
  mutate(
    pct = round(without_IPOS / all * 100, digits = 1)
  ) 

wsa_tbl %>%
  datatable(
    rownames = FALSE,
    colnames = c('CMHSP','Enrollees without IPOS','All Enrollees','Pct without IPOS'),
    caption = 'Percentage of Autism waiver enrollees without IPOS, by CMHSP',
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  ) %>%
  formatStyle(
    'pct',
    background = styleColorBar(data = c(0,100), color = 'grey'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )
  
```

The following analyses will ask different questions for the groups represented here.  

- **Waitlist:** For individuals who do not have an IPOS and are not receiving Autism Waiver services, we will ask whether they receive interim services during their time on the waitlist.  
- **Receiving Services:** For individuals who are receiving autism waiver services, we will ask whether they are receiving services in the amount prescribed in their IPOS and whether their services are being conducted with the appropriate level of observation.

---

# Interim Services for Persons on Autism Waitlist

## Services per Enrollee on Waitlist

For each individual who is autism waiver eligible and on the waitlist, what non-ABA services are he/she receiving while they are waiting? At what frequency?  The table below shows a summary of services received by each enrollee in the waiver who has been placed on the waitlist.^[]  It includes the following information:

- *Services provided:*  A list of services that the person has received
- *Days on waitlist:* The number of days between the enrollee's eligibility date and the most recent service date included in the services data extract (for any enrollee)
- *Days receiving services:*  The number of days between the earliest occurrence of any service following the person's eligibility date and the latest occurrence of any service
- *Percent days covered:* The percentage of the days that the person has been on the waitlist that are overlapped by the provision of services (i.e. `Days receiving services`/`Days on waitlist`)
- *Number of encounters:* The total count of service encounters received.  
- *Contains waiver services:*  This field indicates whether, despite being marked as on the waiting list in WSA, the individual has claims which were submitted as autism waiver services.  If the row is colored red, this indicates that the status recorded in the WSA may require an update.

```{r}

waitlist_svs <-
  wsa %>%
  # Filter individuals who are on waitlist
  filter(
    Status == "Open"
    & IPOSExists == F
    & is.na(Eligibility_Date) == F
  ) %>%
  select(ID,Case_ID,PIHP_CMH_Name,Referral_Date:Eligibility_End_Date) %>%
  left_join(svs, by = c("ID" = "MEDICAID_ID")) %>%
  filter(
    # Only include services occurring after eligibility for autism
    FROM_DATE >= as.Date(Eligibility_Date) 
    # or if there is no service date (to keep people who didn't receive svs)
    | is.na(FROM_DATE) == T 
  ) %>%
  group_by(Case_ID,PIHP_CMH_Name,CPT_CD,SERVICE_DESC,UNIT_TYPE,AUTISM_SRV,Eligibility_Date) %>%
  # Convert units to a common period
  mutate(
    # Recode unit type as numeric conversion factor
    hr_conv = recode(
      as.character(UNIT_TYPE),
      `15 Minutes` = '0.25', 
      `30 Minutes` = '0.5',
      `Hour` = '1', 
      `Up to 15 min` = '0.25',
      `Encounter` = 'NA'
    ),
    svs_conv = recode(
      as.character(SERVICE_DESC),
      `Psychotherapy, 60 min` = 1, 
      `Psychotherapy, 45 min` = 0.75,
      `Psychotherapy, 30 min` = 0.5
    ),
    hr_conv = ifelse(is.na(svs_conv) == F,svs_conv,as.numeric(hr_conv)),
    # Calculate hours
    hrs = UNITS * hr_conv
  ) %>%
  # What is the first and last date of each service?
  mutate(
    first = min(FROM_DATE),
    last = max(FROM_DATE)
  ) %>%
  # What is the number of units received?
  summarize(
    encounters = n(),
    units = sum(UNITS, na.rm = T),
    hrs = sum(hrs, na.rm = T),
    first = min(FROM_DATE),
    last = max(FROM_DATE)
  ) 

waitlist_person <-
  waitlist_svs %>%
  group_by(Case_ID) %>%
  summarize(
    first = min(first),
    last = max(last),
    Eligibility_Date = as.Date(max(Eligibility_Date)),
    svs_list = paste(SERVICE_DESC, collapse = ', '),
    encounters = sum(encounters),
    autism_waiver_svs = sum(AUTISM_SRV, na.rm = T)
  ) %>%
  mutate(
    # Most recent date in services data pulled
    max_date = max(svs$FROM_DATE),
    elig_interval = as.numeric(max_date - Eligibility_Date),
    svs_interval = as.numeric(last - first),
    pct_days_svd = round(svs_interval/elig_interval * 100, digits = 1),
    autism_waiver_svs = autism_waiver_svs > 0
  )

t1 <- waitlist_person %>%
  select(Case_ID,svs_list,elig_interval,svs_interval,pct_days_svd,encounters,autism_waiver_svs) %>%
  ungroup() %>%
  arrange(pct_days_svd)

t2 <- wsa %>%
  # Filter individuals who are on waitlist
  filter(
    Status == "Open"
    & IPOSExists == F
    & is.na(Eligibility_Date) == F
  ) %>%
  select(ID,Case_ID,PIHP_CMH_Name,Referral_Date:Eligibility_End_Date) %>%
  left_join(svs, by = c("ID" = "MEDICAID_ID")) %>%
  filter(
    # Only include services occurring after eligibility for autism
    FROM_DATE >= as.Date(Eligibility_Date) 
    # or if there is no service date (to keep people who didn't receive svs)
    | is.na(FROM_DATE) == T 
  ) %>%
  mutate(
    week = floor_date(FROM_DATE, unit = "week"),
    month = floor_date(FROM_DATE, unit = "month")
  ) %>%
  group_by(Case_ID,month) %>%
  summarize(Encounters = n()) %>%
  ungroup() %>%
  group_by(Case_ID) %>%
  summarise(
    Encounters = paste(Encounters, collapse = ","))

r <- range(t2$Encounters) # define min/max # of encounters for sparkline

t3 <- merge(x = t2, y = t1, by = "Case_ID", all.x = TRUE) # join t1 ad t2


# Define attributes of line spark
line_string <- "type: 'line', lineColor: 'black', fillColor: '#ccc', highlightLineColor: 'orange', highlightSpotColor: 'orange'"

cd <- list(list(targets = 1, render = JS("function(data, type, full){ return '<span class=sparkSeries>' + data + '</span>' }")))

cb = JS(paste0("function (oSettings, json) {
               \n  $('.sparkSeries:not(:has(canvas))').sparkline('html', { ", line_string, " });
               \n}")
        , collapse = "")

dt <- t3 %>%
  datatable(
    data.table(),
    rownames = FALSE,
    colnames = c('WSA ID','Consistency of Encounters',
                 'Services provided',
                 'Days on waitlist','Days receiving services','Percent days covered',
                 'Total Number of encounters','Contains waiver services'),
    caption = 'Service coverage for individuals on Autism Waiver waitlist, by Enrollee',
    extensions = c('Responsive','Buttons'),
    options = list(
      columnDefs = cd,
      pageLength = 5,
      lengthMenu = c(5, 15, 25, 50),
      fnDrawCallback = cb,
      # dom = 't',
      buttons = c('colvis'))
    )

dt$dependencies <- append(dt$dependencies, htmlwidgets:::getDependency("sparkline"))
dt

```


# Conformance with Prescribed Hours for ABA Services

For individuals who are receiving autism waiver services, what is the amount of services and does it match their plan?  The bar chart below shows the percentage of authorized hours of ABA that have been provided for each active autism waiver enrollee over the past 90 days.  Individuals whose services fall outside of the acceptable range (*i.e. <75% or >125%*) of the prescribed hours are differentiated by color.

```{r}

#### Calculate ABA services per week per person ####

# Define ABA codes
aba_cpt <- c("S5108", # Old observation code
             "H2019", # Old treatment codes
             "0364T","0365T","0366T","0367T","0372T","0373T","0374T",
             "0368T","0369T") # Observation codes

aba_rx_hrs <- wsa_ipos %>% select(MEDICAID_ID = Beneficiary_ID,Case_ID,ABA_Hours)

ipos_start <- 
  wsa %>% 
  filter(Status == "Open" & Currently_Inactive == F) %>% 
  select(MEDICAID_ID = ID,IPOS_Start_Date)

last_90 <- as.Date(max(svs$FROM_DATE) - 90)

aba_svs <-
svs %>%
  filter(
    CPT_CD %in% aba_cpt
    & USED_MOD == "U5" # Only include U5 modifier for Autism services
  ) %>%
  droplevels() %>%
  mutate(
    # Recode unit type as numeric conversion factor
    hr_conv = recode(
      as.character(UNIT_TYPE),
      `15 Minutes` = '0.25', 
      `30 Minutes` = '0.5',
      `Hour` = '1', 
      `Up to 15 min` = '0.25',
      `Encounter` = 'NA'
    ),
    hr_conv = as.numeric(hr_conv),
    # Calculate hours
    hrs = UNITS * hr_conv,
    # Create week dates for grouping
    week = floor_date(FROM_DATE, unit = "week")
  ) %>%
  select(PROVIDER_NAME,MEDICAID_ID,CPT_CD:FROM_DATE,week,UNITS,hrs) 

aba_week <-
  aba_svs %>%
  # Get prescribed ABA hours from WSA IPOS
  left_join(aba_rx_hrs, by = "MEDICAID_ID") %>%
  # Get IPOS start date from WSA
  left_join(ipos_start, by = "MEDICAID_ID") %>%
  # Only include services if they occur after the IPOS Start Date
  filter(FROM_DATE >= as.Date(IPOS_Start_Date)) %>%
  # Only include services if they are in most recent 90 days of dataset
  filter(FROM_DATE >= last_90) %>%
  group_by(Case_ID,week) %>%
  summarize(
    hrs = sum(hrs, na.rm = T),
    ABA_Hours = max(ABA_Hours, na.rm = T)
  ) %>%
  mutate(
    # Calculate difference between actual and prescribed hours
    # Negative values indicate hours less than goal
    diff = hrs - ABA_Hours,
    # Calculate % of prescribed hours received
    pct = round(hrs / ABA_Hours * 100, digits = 1),
    # Tag Compliant and Non-Compliant weeks
    in_range_wk = ifelse(between(pct,75,125),T,F)
  ) %>%
  ungroup() %>%
  group_by(Case_ID) %>%
  mutate(
    pct_all = round(sum(hrs) / sum(ABA_Hours) * 100, digits = 1),
    in_range_all = ifelse(between(pct_all,75,125),T,F)
  ) 

aba_per <-
  aba_week %>%
  group_by(Case_ID) %>%
  summarize(
    hrs = sum(hrs, na.rm = T),
    rx_hrs = sum(ABA_Hours, na.rm = T)
  ) %>%
  mutate(
    pct = round(hrs / rx_hrs * 100, digits = 1),
    in_range = ifelse(between(pct,75,125),T,F)
  )

aba_per %>%
  mutate(
    Case_ID = fct_reorder(Case_ID,desc(pct)),
    in_range = ifelse(in_range == T,"Within range","Out of range")
  ) %>%
  plot_ly(x = ~Case_ID, y = ~pct, color = ~in_range, colors = c("#b0145b","#088da5")) %>%
  layout(
    title = "Percent of prescribed hours of ABA services received in last 90 days",
    xaxis = list(title = "Autism Waiver Enrollee",zeroline = F,showline = F,showticklabels = F,showgrid = F), 
    yaxis = list(title = "% of prescribed hours received")
  )


```

The individuals whose ABA service hours are outside of the prescribed range are shown in the table below:

```{r}
aba_per %>% 
  filter(in_range == F) %>% 
  select(-in_range) %>%
  arrange(desc(pct)) %>%
  datatable(
    rownames = FALSE,
    colnames = c('WSA ID','Hours of service','Prescribed Hours','Percent of prescribed hrs'),
    caption = 'Autism waiver enrollees receiving services outside of prescribed range'
  )
```


# Required Observation of ABA Services

According to the Autism Waiver requirements, for every 10 hours of treatment provided (*including codes from the following:* `0364T`:`0367T`,`0372T`:`0374T`,`H2019`) to an individual client, there should be 1 hour of observation by a licensed practitioner(*including codes from the following:* `S5108`,`0368T`,`0369T`).  The table below shows the percentage of active autism waiver clients at each CMHSP who have received at least the required ratio of treatment to observation (10:1) during the most recent 90 days of available service data.  Individuals with fewer than 10 hours of treatment during this period are omitted from the calculations.  If an individual has not been receiving services for 90 days, the calculation applies to the portion of the interval when they have received services.

```{r}
# Note: uses filters defined above for service codes and clients included

observe <-
svs %>%
  filter(
    CPT_CD %in% aba_cpt
    & USED_MOD == "U5" # Only include U5 modifier for Autism services
  ) %>%
  droplevels() %>%
  mutate(
    # Recode unit type as numeric conversion factor
    hr_conv = recode(
      as.character(UNIT_TYPE),
      `15 Minutes` = '0.25', 
      `30 Minutes` = '0.5',
      `Hour` = '1', 
      `Up to 15 min` = '0.25',
      `Encounter` = 'NA'
    ),
    hr_conv = as.numeric(hr_conv),
    # Calculate hours
    hrs = UNITS * hr_conv,
    # Create week dates for grouping
    week = floor_date(FROM_DATE, unit = "week"),
    month = floor_date(FROM_DATE, unit = "month")
  ) %>%
  # Get prescribed ABA hours from WSA IPOS
  left_join(aba_rx_hrs, by = "MEDICAID_ID") %>%
  # Get IPOS start date from WSA
  left_join(ipos_start, by = "MEDICAID_ID") %>%
  # Only include services if they occur after the IPOS Start Date
  filter(week >= IPOS_Start_Date) %>%
  # Only include services if they are in most recent 90 days of dataset
  filter(FROM_DATE >= last_90) %>%
  mutate(
    type = recode(
      CPT_CD,
      `S5108` = "Observation",
      `0368T` = "Observation",
      `0369T` = "Observation",
      `0364T` = "Treatment",
      `0365T` = "Treatment",
      `0366T` = "Treatment",
      `0367T` = "Treatment",
      `0372T` = "Treatment",
      `0373T` = "Treatment",
      `0374T` = "Treatment",
      `H2019` = "Treatment"
    )
  ) %>%
  select(
    Case_ID,PROVIDER_NAME,type,CPT_CD,SERVICE_DESC,
    month,week,FROM_DATE,UNITS,hrs
  ) %>%
  group_by(Case_ID,PROVIDER_NAME,type) %>%
  summarize(hrs = sum(hrs)) %>%
  ungroup() %>%
  spread(type, hrs) %>%
  # Exclude cases with < 10 hours of treatment
  filter(Treatment >= 10) %>%
  mutate(
    obs_ratio = round(Observation / Treatment, digits = 2),
    meets_req = obs_ratio >= 0.1 & obs_ratio <= 1.0
  ) 

observe_agg <-
observe %>%
  group_by(PROVIDER_NAME) %>%
  summarize(
    meets_req = sum(meets_req, na.rm = T),
    n = n()
  ) %>%
  mutate(pct = round(meets_req / n * 100, digits = 1)) 
  
observe_agg %>%  
  datatable(
    rownames = FALSE,
    colnames = c('CMHSP','Enrollees with required observation','All Enrollees','Percent meeting standard'),
    caption = 'Percentage of Autism waiver enrollees receiving required observation, by CMHSP',
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  ) %>%
  formatStyle(
    'pct',
    background = styleColorBar(data = c(0,100), color = 'grey'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )

```

For a detailed view of the individuals who have not received the appropriate level of observation for ABA services, please see the table below:

```{r}


observe %>%
  filter(meets_req == F) %>%
  select(-meets_req) %>%
  arrange(desc(obs_ratio)) %>%
  datatable(
    rownames = FALSE,
    colnames = c('WSA ID','CMHSP','Hours of treatment','Hours of observation','Ratio of observation to treatment'),
    caption = 'Individuals without required observation for ABA treatment',
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  )

```


The individuals whose ABA service hours are outside of the prescribed range are shown in the table below:

```{r}
aba_per %>% 
  filter(in_range == F) %>% 
  select(-in_range) %>%
  arrange(desc(pct)) %>%
  datatable(
    rownames = FALSE,
    colnames = c('Enrollee ID','Hours of service','Prescribed Hours','Percent of prescribed hrs'),
    caption = 'Autism waiver enrollees receiving services outside of prescribed range'
  )
```


# Required Observation of ABA Services

According to the Autism Waiver requirements, for every 10 hours of treatment provided (*including codes from the following:* `0364T`:`0367T`,`0372T`:`0374T`,`H2019`) to an individual client, there should be 1 hour of observation by a licensed practitioner(*including codes from the following:* `S5108`,`0368T`,`0369T`).  The table below shows the percentage of active autism waiver clients at each CMHSP who have received at least the required ratio of treatment to observation (10:1) during the most recent 90 days of available service data.  Individuals with fewer than 10 hours of treatment during this period are omitted from the calculations.  If an individual has not been receiving services for 90 days, the calculation applies to the portion of the interval when they have received services.

```{r}
# Note: uses filters defined above for service codes and clients included

observe <-
svs %>%
  filter(
    CPT_CD %in% aba_cpt
    & USED_MOD == "U5" # Only include U5 modifier for Autism services
  ) %>%
  droplevels() %>%
  mutate(
    # Recode unit type as numeric conversion factor
    hr_conv = recode(
      as.character(UNIT_TYPE),
      `15 Minutes` = '0.25', 
      `30 Minutes` = '0.5',
      `Hour` = '1', 
      `Up to 15 min` = '0.25',
      `Encounter` = 'NA'
    ),
    hr_conv = as.numeric(hr_conv),
    # Calculate hours
    hrs = UNITS * hr_conv,
    # Create week dates for grouping
    week = floor_date(FROM_DATE, unit = "week"),
    month = floor_date(FROM_DATE, unit = "month")
  ) %>%
  # Get prescribed ABA hours from WSA IPOS
  left_join(aba_rx_hrs, by = "MEDICAID_ID") %>%
  # Get IPOS start date from WSA
  left_join(ipos_start, by = "MEDICAID_ID") %>%
  # Only include services if they occur after the IPOS Start Date
  filter(week >= IPOS_Start_Date) %>%
  # Only include services if they are in most recent 90 days of dataset
  filter(FROM_DATE >= last_90) %>%
  mutate(
    type = recode(
      CPT_CD,
      `S5108` = "Observation",
      `0368T` = "Observation",
      `0369T` = "Observation",
      `0364T` = "Treatment",
      `0365T` = "Treatment",
      `0366T` = "Treatment",
      `0367T` = "Treatment",
      `0372T` = "Treatment",
      `0373T` = "Treatment",
      `0374T` = "Treatment",
      `H2019` = "Treatment"
    )
  ) %>%
  select(
    Case_ID,PROVIDER_NAME,type,CPT_CD,SERVICE_DESC,
    month,week,FROM_DATE,UNITS,hrs
  ) %>%
  group_by(Case_ID,PROVIDER_NAME,type) %>%
  summarize(hrs = sum(hrs)) %>%
  ungroup() %>%
  spread(type, hrs) %>%
  # Exclude cases with < 10 hours of treatment
  filter(Treatment >= 10) %>%
  mutate(
    obs_ratio = round(Observation / Treatment, digits = 2),
    meets_req = obs_ratio >= 0.1 & obs_ratio <= 1.0
  ) 

observe_agg <-
observe %>%
  group_by(PROVIDER_NAME) %>%
  summarize(
    meets_req = sum(meets_req, na.rm = T),
    n = n()
  ) %>%
  mutate(pct = round(meets_req / n * 100, digits = 1)) 
  
observe_agg %>%  
  datatable(
    rownames = FALSE,
    colnames = c('CMHSP','Enrollees with required observation','All Enrollees','Percent meeting standard'),
    caption = 'Percentage of Autism waiver enrollees receiving required observation, by CMHSP',
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  ) %>%
  formatStyle(
    'pct',
    background = styleColorBar(c(0,100), 'grey'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )

```

For a detailed view of the individuals who have not received the appropriate level of observation for ABA services, please see the table below:

```{r}

observe %>%
  filter(meets_req == FALSE) %>%
  select(-meets_req) %>%
  arrange(desc(obs_ratio)) %>%
  datatable(
    rownames = FALSE,
    colnames = c('WSA ID','CMHSP','Treatment Hours','Observation Hours','Observation Ratio'),
    caption = 'Autism waiver enrollees not receiving required observation',
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  )

```

