---
title: "Autism Benefit Compliance"
output:
  html_document:
    self_contained: yes
    theme: yeti
  word_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load_libs}
# Load required packages
library(tidyverse); library(stringr); library(magrittr); library(lubridate); library(readxl)
library(DT); library(forcats); library(plotly); library(sparkline);library(feather);library(webshot)

chart_color <- c("#e34a33")
```

```{r get_data}

# Read in .csv files as dataframes
csv_path <- "data/"
svs <- read_feather(paste0(csv_path,"svs.feather"))
# Read in .csv files as dataframes
wsa <- read_feather(paste0(csv_path,"wsa.feather"))
# If re-opened, client is given new 'Case ID' but Medicaid ID (i.e. 'ID') should be same

# Read in data related to authorized units
wsa_ipos <- read_feather(paste0(csv_path,"wsa_ipos.feather"))

```

<h4>This *summary* report was generated for Region 10 PIHP on `r Sys.Date()` using data from `r min(svs$FROM_DATE, na.rm = T)` - `r max(svs$FROM_DATE, na.rm = T)`</h4>

*Please note that the data in this report may not reflect all services that enrollees received from the Autism Benefit.  Specifically, services may not appear for more recent periods if claims have not yet been submitted.*

---

# Overdue Re-Evaluations

```{r reeval_create}
reeval_tbl <-
wsa %>%
  filter(Status == "Open") %>%
  mutate(
    past_due = ReEvaluation_Due_Date < today()
  ) %>%
  group_by(PIHP_CMH_Name) %>%
  summarize(
    numerator = sum(past_due, na.rm = T),
    denominator = n()
  ) %>%
  mutate(
    pct = round(numerator / denominator * 100, digits = 1)
  ) %>%
  arrange(as.character(PIHP_CMH_Name)) %>%
  # Add subtotal line for PIHP
  rbind(
    ., 
    data.frame(
      PIHP_CMH_Name = "Region 10", 
      t(colSums(.[2:3])),
      pct = round(colSums(.[2])/colSums(.[3])*100,digits = 1)
    )
  ) %>%
  select(PIHP_CMH_Name,denominator,numerator,pct)
```

```{r reeval_table}
reeval_tbl %>%
  datatable(
    rownames = FALSE,
    colnames = c('Organization','Total Enrollees','Overdue Re-Evaluations','% of Enrollees Overdue'),
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  ) %>%
  formatStyle(
    'pct',
    background = styleColorBar(c(0,100), '#fc8d59'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )
```
<br>
```{r reeval_chart}
reeval_tbl %>%
  mutate(
    PIHP_CMH_Name = fct_relevel(
      PIHP_CMH_Name,
      c(
        "Genesee Health System","Lapeer County CMH","Sanilac County CMH",
        "St. Clair County CMH","Region 10"
      )
    )
  ) %>%
  plot_ly(
    x = ~PIHP_CMH_Name,
    y = ~pct,
    color = I(chart_color)
  ) %>%
  add_bars(
    hoverinfo = 'text',
    text = ~paste0(
      "At ", PIHP_CMH_Name,", ", numerator, " out of ",denominator," people (",pct,"%)<br>",
      "are overdue to receive re-evaluations."
    )
  ) %>%
  layout(
    title = "Compliance with Re-Evaluations",
    xaxis = list(title = "", tickangle = 45),
    yaxis = list(
      title = "% of total enrollees with overdue re-evaluations",
      range = c(0, 100)
    ),
    margin = list(l = 60, r = 30, b = 150, t = 40, pad = 4)
  )
```

---

# Overdue Plans of Service

```{r bpos_create}
bpos_tbl <-
wsa %>%
  filter(Status == "Open" & IPOSExists == T) %>%
  mutate(
    past_due = IPOS_Due_Date < today()
  ) %>%
  group_by(PIHP_CMH_Name) %>%
  summarize(
    numerator = sum(past_due, na.rm = T),
    denominator = n()
  ) %>%
  mutate(
    pct = round(numerator / denominator * 100, digits = 1)
  ) %>%
  arrange(as.character(PIHP_CMH_Name)) %>%
  # Add subtotal line for PIHP
  rbind(
    ., 
    data.frame(
      PIHP_CMH_Name = "Region 10", 
      t(colSums(.[2:3])),
      pct = round(colSums(.[2])/colSums(.[3])*100,digits = 1)
    )
  ) %>%
  select(PIHP_CMH_Name,denominator,numerator,pct)
```

```{r bpos_table}
bpos_tbl %>%
  datatable(
    rownames = FALSE,
    colnames = c('Organization','Total Enrollees with BPOS','Overdue BPOS','% of Enrollees Overdue'),
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  ) %>%
  formatStyle(
    'pct',
    background = styleColorBar(c(0,100), '#fc8d59'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )
```
<br>
```{r bpos_chart}
bpos_tbl %>%
  mutate(
    PIHP_CMH_Name = fct_relevel(
      PIHP_CMH_Name,
      c(
        "Genesee Health System","Lapeer County CMH","Sanilac County CMH",
        "St. Clair County CMH","Region 10"
      )
    )
  ) %>%
  plot_ly(
    x = ~PIHP_CMH_Name,
    y = ~pct,
    color = I(chart_color)
  ) %>%
  add_bars(
    hoverinfo = 'text',
    text = ~paste0(
      "At ", PIHP_CMH_Name,", ", numerator, " out of ",denominator," people (",pct,"%)<br>",
      "with a BPOS are overdue to receive an updated BPOS."
    )
  ) %>%
  layout(
    title = "Compliance with Updated Plans of Service",
    xaxis = list(title = "", tickangle = 45),
    yaxis = list(
      title = "% of total enrollees with overdue BPOS",
      range = c(0, 100)
    ),
    margin = list(l = 60, r = 30, b = 150, t = 40, pad = 4)
  )
```

---

# Enrollees Without Behavioral Plan of Service

Enrollees approved for ABA and are awaiting the development of an ABA behavior treatment plan to begin services.

```{r start_create}
wsa_tbl <-
wsa %>%
  filter(Status == "Open") %>%
  group_by(PIHP_CMH_Name) %>%
  summarize(
    numerator = n() - sum(IPOSExists, na.rm = T),
    denominator = n()
  ) %>%
  mutate(
    pct = round(numerator / denominator * 100, digits = 1)
  ) %>%
  arrange(as.character(PIHP_CMH_Name)) %>%
  # Add subtotal line for PIHP
  rbind(
    ., 
    data.frame(
      PIHP_CMH_Name = "Region 10", 
      t(colSums(.[2:3])),
      pct = round(colSums(.[2])/colSums(.[3])*100,digits = 1)
    )
  ) %>%
  select(PIHP_CMH_Name,denominator,numerator,pct)
```

```{r start_table}
wsa_tbl %>%
  datatable(
    rownames = FALSE,
    colnames = c('Organization','Total Enrollees','Total Enrollees without a BPOS','% of Enrollees without BPOS'),
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  ) %>%
  formatStyle(
    'pct',
    background = styleColorBar(c(0,100), '#fc8d59'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )
```
<br>
```{r start_chart}
wsa_tbl %>%
  mutate(
    PIHP_CMH_Name = fct_relevel(
      PIHP_CMH_Name,
      c(
        "Genesee Health System","Lapeer County CMH","Sanilac County CMH",
        "St. Clair County CMH","Region 10"
      )
    )
  ) %>%
  plot_ly(
    x = ~PIHP_CMH_Name,
    y = ~pct,
    color = I(chart_color)
  ) %>%
  add_bars(
    hoverinfo = 'text',
    text = ~paste0(
      "At ", PIHP_CMH_Name,", ", numerator, " out of ",denominator," people (",pct,"%)<br>",
      "were overdue to start autism services."
    )
  ) %>%
  layout(
    title = "Overdue Start Date",
    xaxis = list(title = "", tickangle = 45),
    yaxis = list(
      title = "% of total enrollees with overdue service start",
      range = c(0, 100)
    ),
    margin = list(l = 60, r = 30, b = 150, t = 40, pad = 4)
  )
  

```

---

# Range of Intensity Non-Compliance

What percent of beneficiaries have received ABA services which fall outside of the acceptable range (*i.e. <75% or >125%*) of the prescribed hours?

```{r roi_create}

#### Calculate ABA services per week per person 

# Define ABA codes
aba_cpt <- c("S5108", # Old observation code
             "H2019", # Old treatment codes
             "0364T","0365T","0366T","0367T","0372T","0373T","0374T",
             "0368T","0369T") # Observation codes

aba_rx_hrs <- wsa_ipos %>% select(MEDICAID_ID = Beneficiary_ID,Case_ID,ABA_Hours)

ipos_start <- 
  wsa %>% 
  filter(Status == "Open" & Currently_Inactive == F) %>% 
  select(MEDICAID_ID = ID,IPOS_Start_Date)

last_90 <- as.Date(max(svs$FROM_DATE, na.rm = T) - 90)

aba_svs <-
svs %>%
  filter(
    CPT_CD %in% aba_cpt
    & USED_MOD == "U5" # Only include U5 modifier for Autism services
  ) %>%
  # Now remove observation codes
  filter(!CPT_CD %in% c("S5108","0368T","0369T")) %>%
  droplevels() %>%
  mutate(
    # Recode unit type as numeric conversion factor
    hr_conv = recode(
      as.character(UNIT_TYPE),
      `15 Minutes` = '0.25', 
      `30 Minutes` = '0.5',
      `Hour` = '1', 
      `Up to 15 min` = '0.25',
      `Encounter` = 'NA'
    ),
    hr_conv = as.numeric(hr_conv),
    # Calculate hours
    hrs = UNITS * hr_conv,
    # Create week dates for grouping
    week = floor_date(FROM_DATE, unit = "week")
  ) %>%
  select(PROVIDER_NAME,MEDICAID_ID,CPT_CD:FROM_DATE,week,UNITS,hrs) 

aba_summary <-
  aba_svs %>%
  # Get prescribed ABA hours from WSA IPOS
  left_join(aba_rx_hrs, by = "MEDICAID_ID") %>%
  # Get IPOS start date from WSA
  left_join(ipos_start, by = "MEDICAID_ID") %>%
  # Only include services if they occur after the IPOS Start Date
  filter(FROM_DATE >= IPOS_Start_Date) %>%
  # Only include services if they are in most recent 90 days of dataset
  filter(FROM_DATE >= last_90) %>%
  group_by(Case_ID,PROVIDER_NAME,week) %>%
  summarize(
    hrs = sum(hrs, na.rm = T),
    ABA_Hours = max(ABA_Hours)
  ) %>%
  mutate(
    # Calculate difference between actual and prescribed hours
    # Negative values indicate hours less than goal
    diff = hrs - ABA_Hours,
    # Calculate % of prescribed hours received
    # Tag Compliant and Non-Compliant weeks
   
    pct_all = round(sum(hrs) / sum(ABA_Hours) * 100, digits = 1),
    in_range_all = ifelse(between(pct_all,75,125),T,F)
  ) %>%
  ungroup() %>%
  group_by(Case_ID,PROVIDER_NAME) %>%
  summarise(
    hrs = sum(hrs, na.rm = T),
    ABA_Hours = sum(ABA_Hours, na.rm = T)
  ) %>%
  mutate(
    pct = round(hrs / ABA_Hours * 100, digits = 1),
    in_range = ifelse(between(pct,75,125),T,F)
  ) %>%
  group_by(PROVIDER_NAME) %>%
  summarise(
    in_range =  sum(in_range),
    denominator = n()
  ) %>%
  mutate(
    numerator = denominator - in_range,
    pct = round(numerator/denominator * 100, digits = 1)
  ) %>% 
  select(PROVIDER_NAME,numerator,denominator,pct) %>%
  arrange(as.character(PROVIDER_NAME)) %>%
  # Add subtotal line for PIHP
  rbind(
    ., 
    data.frame(
      PROVIDER_NAME = "Region 10", 
      t(colSums(.[2:3])),
      pct = round(colSums(.[2])/colSums(.[3])*100,digits = 1)
    )
  ) %>%
  select(PROVIDER_NAME,denominator,numerator,pct)

```

```{r roi_table}
aba_summary %>%
  datatable(
    rownames = FALSE,
    colnames = c('Organization','Total Enrollees with a Behavioral Plan of Service','Enrollees Non-Compliant','% Non-Compliant'),
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  ) %>%
  formatStyle(
    'pct',
    backgroundSize = '100% 90%',
    background = styleColorBar(c(0,100), '#fc8d59'),
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )
```
<br>
```{r roi_chart}
aba_summary %>%
  mutate(
    PROVIDER_NAME = fct_relevel(
      PROVIDER_NAME,
      c(
        "Genesee Health System","Lapeer County CMH","Sanilac County CMH",
        "St. Clair County CMH","Region 10"
      )
    )
  ) %>%
  plot_ly(
    x = ~PROVIDER_NAME,
    y = ~pct,
    color = I(chart_color)
  ) %>%
  add_bars(
    hoverinfo = 'text',
    text = ~paste0(
      "At ", PROVIDER_NAME,", ", numerator, " out of ",denominator," people (",pct,"%)<br>",
      "did not receive ABA services at the required intensity."
    )
  ) %>%
  layout(
    title = "Range of Intensity Non-Compliance",
    xaxis = list(title = "", tickangle = 45),
    yaxis = list(
      title = "% non-compliant",
      range = c(0, 100)
    ),
    margin = list(l = 60, r = 30, b = 150, t = 40, pad = 4)
  )
```


---

# Observation Non-Compliance

```{r observe_create}
# Note: uses filters defined above for service codes and clients included

observe <-
svs %>%
  filter(
    CPT_CD %in% aba_cpt
    & USED_MOD == "U5" # Only include U5 modifier for Autism services
  ) %>%
  droplevels() %>%
  mutate(
    # Recode unit type as numeric conversion factor
    hr_conv = recode(
      as.character(UNIT_TYPE),
      `15 Minutes` = '0.25', 
      `30 Minutes` = '0.5',
      `Hour` = '1', 
      `Up to 15 min` = '0.25',
      `Encounter` = 'NA'
    ),
    hr_conv = as.numeric(hr_conv),
    # Calculate hours
    hrs = UNITS * hr_conv,
    # Create week dates for grouping
    week = floor_date(FROM_DATE, unit = "week"),
    month = floor_date(FROM_DATE, unit = "month")
  ) %>%
  # Get prescribed ABA hours from WSA IPOS
  left_join(aba_rx_hrs, by = "MEDICAID_ID") %>%
  # Get IPOS start date from WSA
  left_join(ipos_start, by = "MEDICAID_ID") %>%
  # Only include services if they occur after the IPOS Start Date
  filter(week >= IPOS_Start_Date) %>%
  # Only include services if they are in most recent 90 days of dataset
  filter(FROM_DATE >= last_90) %>%
  mutate(
    type = recode(
      CPT_CD,
      `S5108` = "Observation",
      `0368T` = "Observation",
      `0369T` = "Observation",
      `0364T` = "Treatment",
      `0365T` = "Treatment",
      `0366T` = "Treatment",
      `0367T` = "Treatment",
      `0372T` = "Treatment",
      `0373T` = "Treatment",
      `0374T` = "Treatment",
      `H2019` = "Treatment"
    )
  ) %>%
  select(
    Case_ID,PROVIDER_NAME,type,CPT_CD,SERVICE_DESC,
    month,week,FROM_DATE,UNITS,hrs
  ) %>%
  group_by(Case_ID,PROVIDER_NAME,type) %>%
  summarize(hrs = sum(hrs)) %>%
  ungroup() %>%
  mutate(hrs = ifelse(is.na(hrs),0,hrs)) %>%
  spread(type, hrs) %>%
  mutate(
    obs_ratio = round(Observation / Treatment, digits = 2),
    meets_req = obs_ratio >= 0.1 & obs_ratio <= 1.0
  ) %>%
  # Exclude cases with < 10 hours of treatment if they don't meet criteria
  filter(Treatment >= 10 | (Treatment < 10 & meets_req == T))

observe_agg <-
observe %>%
  group_by(PROVIDER_NAME) %>%
  summarize(
    meets_req = sum(meets_req, na.rm = T),
    denominator = n()
  ) %>%
  mutate(
    numerator = denominator - meets_req,
    pct = round(numerator / denominator * 100, digits = 1)
  ) %>% 
  select(PROVIDER_NAME,numerator,denominator,pct) %>%
  arrange(as.character(PROVIDER_NAME)) %>%
  # Add subtotal line for PIHP
  rbind(
    ., 
    data.frame(
      PROVIDER_NAME = "Region 10", 
      t(colSums(.[2:3])),
      pct = round(colSums(.[2])/colSums(.[3])*100,digits = 1)
    )
  ) %>%
  select(PROVIDER_NAME,denominator,numerator,pct)

```

```{r observe_table}
observe_agg %>%
  datatable(
    rownames = FALSE,
    colnames = c('Organization','Enrollees with 10+ Hrs of ABA Svs','Enrollees Non-Compliant','% Non-Compliant'),
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  ) %>%
  formatStyle(
    'pct',
    background = styleColorBar(c(0,100), '#fc8d59'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )
```
<br>
```{r observe_chart}
observe_agg %>%
  mutate(
    PROVIDER_NAME = fct_relevel(
      PROVIDER_NAME,
      c(
        "Genesee Health System","Lapeer County CMH","Sanilac County CMH",
        "St. Clair County CMH","Region 10"
      )
    )
  ) %>%
  plot_ly(
    x = ~PROVIDER_NAME,
    y = ~pct,
    color = I(chart_color)
  ) %>%
  add_bars(
    hoverinfo = 'text',
    text = ~paste0(
      "At ", PROVIDER_NAME,", ", numerator, " out of ",denominator," people (",pct,"%)<br>",
      "did not receive the required level of observation for ABA services."
    )
  ) %>%
  layout(
    title = "Observation Non-Compliance",
    xaxis = list(title = "", tickangle = 45),
    yaxis = list(
      title = "% non-compliant",
      range = c(0, 100)
    ),
    margin = list(l = 60, r = 30, b = 150, t = 40, pad = 4)
  )
```


---

# Glossary

**ABA Services:** Services provided by the CMHSP, or contracted ABA provider, which include the S5108, H2019, 0364T, 0365T, 0366T, 0367T, 0372T, 0373T, 0374T, 0368T, and 0369T service codes reported with a U5 modifier. (Note: U5 modifier must be reported on all encounters covered by the EPSDT Autism Benefit).

**Behavioral Plan of Service (BPOS):** In WSA, this includes the ABA behavioral plan and the related Individual Plan of Service created by the CMHSP. 

**Days Without a Behavioral Plan of Service: **  The number of days without a Behavioral Plan of Service following eligibility determination date.  

**Eligibility Determination Date:** The date that MDHHS determines that an individual is eligible for ABA services.  

**Enrollee:** A person found eligible for the Autism Benefit, as determined by MDHHS. 

**Family Training:**  Required service provided by qualified practitioner. At least 1 encounter / 90 days meets the standard.  The service codes for Family Training are 0370T and 0371T.

**Hours of Service: ** ABA services including supervision/observation. 

**Hours of Treatment:** ABA services excluding supervision/observation.

**Number of Encounters:** The total count of service encounters received. Note: if multiple units are received as part of a single encounter, the encounter will only be counted once.    

**Observation of ABA Services:** The standard is for every 10 hours of ABA treatment provided to an individual enrollee, 1 hour of supervision/observation by a qualified practitioner will be provided. The service codes for observation are S5108, 0368T, and 0369T with a U5 modifier.

**Overdue Plans of Service: ** Cases with an expired plan of service in WSA.

**Overdue Re-Evaluation:** Cases with an expired re-evaluation (364 days from most recent evaluation end date) in WSA.

**Overdue Service Start:** Individuals determined eligible for ABA services awaiting the development of an ABA treatment plan to begin services.

**Prescribed Hours:** The number of hours an enrollee is authorized for ABA services based on assessed need and recommendation. This number is entered in WSA by the CMHSPs.

**Range of Intensity:** The hours of ABA treatment provided; within 25% +/- of prescribed hours meets the standard.

**Ratio of Observation to Treatment:** The result of dividing the number of hours of 'observation of ABA Services' by 'hours of treatment'. A ratio value of greater than or equal to 0.1 (1 hour of observation for every 10 hours of treatment) meets the standard.  Note that this metric only includes enrollees with a BPOS who received >= 10 hours of treatment during the reporting period.

**Services Provided:** A list of services (i.e. the descriptions which accompany HCPCS codes) that the enrollee has received. 

**Waiver Support Application (WSA):** The electronic platform used by MDHHS, PIHPs, and CMHs to monitor the Autism Benefit and enrollees. 



