---
title: "Autism Service Compliance"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

**Please note that this is currently a draft report under development with analysts at Region 10 PIHP**

```{r}
# Install packages if you need them
# if (!require("tidyverse")) install.packages("tidyverse")
# if (!require("stringr")) install.packages("stringr")
# if (!require("magrittr")) install.packages("magrittr")
# if (!require("lubridate")) install.packages("lubridate")
# if (!require("plotly")) install.packages("plotly")
# if (!require("readxl")) install.packages("readxl")

# Load required packages
library(tidyverse); library(stringr); library(magrittr); library(lubridate); library(readxl)
library(DT); library(forcats); library(plotly)
```


# Reading in Data

The data for these analyses is drawn from two primary sources: 

- *Service data* from the region's MIX EMR system and 
- *Waiver enrollment* reports generated from the MDHHS WSA system.

First, we read in the services data and process it for analysis.  Note that we are currently reading from .csv file output for this draft analysis.  A repeatable, automated process will rely on connection to database tables updated via ETL.

```{r}
# Read in .csv files as dataframes
csv_path <- "C:/Users/joshh/OneDrive - TBD Solutions LLC/files/Region10/Autism/"
genesee <- read_csv(paste0(csv_path,"Autism.Genesse.csv"), skip = 7)
lapeer <- read_csv(paste0(csv_path,"Autism.Lapeer.csv"), skip = 7)
sanilac <- read_csv(paste0(csv_path,"Autism.Sanilac.csv"), skip = 7)
stclair <- read_csv(paste0(csv_path,"Autism.StClair.csv"), skip = 7)

# Bind separate CMH dataframes together
svs <- genesee %>% bind_rows(lapeer,sanilac,stclair)

# Remove separate files from environment
rm(csv_path); rm(genesee); rm(lapeer); rm(sanilac); rm(stclair)

# Remove cols where all values are NA
svs <- Filter(function(x)!all(is.na(x)), svs)

# Clean colnames (rm spaces, other symbols, add underscore sep)
names(svs) <- gsub(":", "", names(svs))
names(svs) <- gsub(" |-", "_", names(svs))

# Clean data to prepare for analysis
svs <-
svs %>%
  # Clean Medicaid ID field
  mutate(
    # Trim lead / trail whitespace
    MEDICAID_ID = str_trim(MEDICAID_ID),
    # Remove alpha and special chars
    MEDICAID_ID = str_replace_all(MEDICAID_ID, "[[:alpha:]]", ""),
    MEDICAID_ID = str_replace_all(MEDICAID_ID, "[[:punct:]]", ""),
    # Convert blanks to NA
    MEDICAID_ID = ifelse(MEDICAID_ID == "", yes = NA, no = MEDICAID_ID), 
    # If string > 10 chars, include only last 10 chars
    MEDICAID_ID = ifelse(nchar(as.character(MEDICAID_ID)) > 10,
                      yes = substr(MEDICAID_ID, 
                                   start = nchar(as.character(MEDICAID_ID)) - 9, 
                                   stop = nchar(as.character(MEDICAID_ID))),
                      no = MEDICAID_ID),
    # If string < 10 chars, pad with leading zeroes
    MEDICAID_ID = ifelse(nchar(as.character(MEDICAID_ID)) < 10,
                      yes = sprintf("%010d", as.integer(MEDICAID_ID)),
                      no = MEDICAID_ID),
    # Make 'NA' & 0000000000 to NA
    MEDICAID_ID = ifelse(MEDICAID_ID %in% c("        NA","NA","0000000000"), 
                      yes = NA,
                      no = MEDICAID_ID),
    # Convert to factor
    MEDICAID_ID = as.factor(MEDICAID_ID)
  ) %>%
  # Change numeric ID vars to characters
  mutate_at(
    .cols = vars(PRV_ID,CON_ID,CLM_ID:PRIM_INS_ID),
    .funs = funs(as.character)
  ) %>%
  # Change all character columns to factors
  mutate_if(is.character,as.factor) %>%
  # Clean date fields to prepare for analysis
  mutate(
    FROM_DATE = mdy(FROM_DATE),
    THRU_DATE = mdy(THRU_DATE)
  ) %>%
  # Transform Y/N responses into logical vars
  mutate_at(
    .cols = vars(HAB_WAIVER:HMP_BUCKET),
    .funs = funs(. == "Y")
  ) 
```

Next, we read in the waiver enrollment data and process it:

```{r}
# Read in .csv files as dataframes
csv_path <- "C:/Users/joshh/OneDrive - TBD Solutions LLC/files/Region10/Autism/"
wsa <- read_excel(paste0(csv_path,"WSA All Cases All Data 05-30-17.xls"), skip = 1)

# Clean colnames (rm spaces, other symbols, add underscore sep)
names(wsa) <- gsub(" |-|/", "_", names(wsa))
names(wsa) <- gsub("[[:space:]]", "", names(wsa))

wsa %<>%
  mutate(
    Case_ID = as.character(Case_ID)
  ) %>%
  mutate_at(
    .cols = vars(Days_Bet_Ref_Eval,Days_Bet_Elig_IPOS),
    .funs = funs(as.numeric)
  ) %>%
  # Change all character columns to factors
  mutate_if(is.character,as.factor) %>%
  # Transform Y/N responses into logical vars
  mutate_at(
    .cols = vars(
      IPOSExists,
      Telepractice_Authorization_Requested,
      Currently_Inactive,
      Has_Past_Inactive
    ),
    .funs = funs(. == "Yes")
  )
  
# If closed and opened, client is given a new 'Case ID' 
# but Medicaid ID (i.e. 'ID') should be the same

# Read in data related to authorized units

wsa_ipos <- read_excel(paste0(csv_path,"06-08-17 WSA IPOS Data Report.xls"), skip = 1)

# Clean colnames (rm spaces, other symbols, add underscore sep)
names(wsa_ipos) <- gsub(" |-|/", "_", names(wsa_ipos))
names(wsa_ipos) <- gsub("[[:space:]]", "", names(wsa_ipos))

wsa_ipos %<>%
  mutate(
    Case_ID = as.character(Case_ID)
  ) %>%
  mutate_at(
    .cols = vars(Days_Without_IPOS,Next_IPOS_Due_In,ABA_Hours),
    .funs = funs(as.numeric)
  ) %>%
  # Change all character columns to factors
  mutate_if(is.character,as.factor) %>%
  # Transform Y/N responses into logical vars
  mutate_at(
    .cols = vars(IPOS_Exists),
    .funs = funs(. == "Yes")
  )
```

---

# Eligible Autism Waiver Enrollees Receiving services

As an initial summary, we look at the percent of persons who are enrolled in the autism waiver (i.e. `Status == "Open"`), but who do not have an Individualized Plan of Service (IPOS).

```{r}
wsa_tbl <-
wsa %>%
  filter(Status == "Open") %>%
  group_by(PIHP_CMH_Name) %>%
  summarize(
    without_IPOS = n() - sum(IPOSExists, na.rm = T),
    all = n()
  ) %>%
  mutate(
    pct = round(without_IPOS / all * 100, digits = 1)
  ) 

wsa_tbl %>%
  datatable(
    rownames = FALSE,
    colnames = c('CMHSP','Enrollees without IPOS','All Enrollees','Pct without IPOS'),
    caption = 'Percentage of Autism waiver enrollees without IPOS, by CMHSP',
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  ) %>%
  formatStyle(
    'pct',
    background = styleColorBar(wsa_tbl$pct, 'grey'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )
  
```

The following analyses will ask different questions for the groups represented here.  

- **Waitlist:** For individuals who do not have an IPOS and are not receiving Autism Waiver services, we will ask whether they receive interim services during their time on the waitlist.  
- **Receiving Services:** For individuals who are receiving autism waiver services, we will ask whether they are receiving services in the amount prescribed in their IPOS and whether their services are being conducted with the appropriate level of observation.

---

# Interim Services for Persons on Autism Waitlist

## Services per Enrollee on Waitlist

For each individual who is autism waiver eligible and on the waitlist, what services are he/she receiving while they are waiting? At what frequency? 

```{r}
waitlist_svs <-
  wsa %>%
  # Filter individuals who are on waitlist
  # Is this an accurate way to get at these individuals?
  filter(
    Status == "Open"
    & IPOSExists == F
    & is.na(Eligibility_Date) == F
  ) %>%
  select(ID,PIHP_CMH_Name,Referral_Date:Eligibility_End_Date) %>%
  left_join(svs, by = c("ID" = "MEDICAID_ID")) %>%
  filter(
    # Only include services occurring after eligibility for autism
    FROM_DATE >= as.Date(Eligibility_Date) 
    # or if there is no service date (to keep people who didn't receive svs)
    | is.na(FROM_DATE) == T 
  ) %>%
  group_by(ID,PIHP_CMH_Name,CPT_CD,SERVICE_DESC,UNIT_TYPE,AUTISM_SRV,Eligibility_Date) %>%
  # Convert units to a common period
  mutate(
    # Recode unit type as numeric conversion factor
    hr_conv = recode(
      as.character(UNIT_TYPE),
      `15 Minutes` = '0.25', 
      `30 Minutes` = '0.5',
      `Hour` = '1', 
      `Up to 15 min` = '0.25',
      `Encounter` = 'NA'
    ),
    svs_conv = recode(
      as.character(SERVICE_DESC),
      `Psychotherapy, 60 min` = 1, 
      `Psychotherapy, 45 min` = 0.75,
      `Psychotherapy, 30 min` = 0.5
    ),
    hr_conv = ifelse(is.na(svs_conv) == F,svs_conv,as.numeric(hr_conv)),
    # Calculate hours
    hrs = UNITS * hr_conv
  ) %>%
  # What is the first and last date of each service?
  mutate(
    first = min(FROM_DATE),
    last = max(FROM_DATE)
  ) %>%
  # What is the number of units received?
  summarize(
    units = sum(UNITS, na.rm = T),
    hrs = sum(hrs, na.rm = T),
    first = min(FROM_DATE),
    last = max(FROM_DATE)
  )
```

## Summary of Services per Enrollee

The table below shows a summary of services received by each enrollee in the waiver who has been placed on the waitlist.  It includes the following information:

- *Services provided:*  A list of services that the person has received
- *Days on waitlist:* The number of days between the enrollee's eligibility date and the most recent service date included in the services data extract (for any enrollee)
- *Days receiving services:*  The number of days between the earliest occurrence of any service following the person's eligibility date and the latest occurrence of any service
- *Percent days covered:* The percentage of the days that the person has been on the waitlist that are overlapped by the provision of services (i.e. `Days receiving services`/`Days on waitlist`)
- *Approximate hours:* The total number of hours of services received of any service that can be converted to a standard hour measurement.  This does not include units with the unit type of `Encounter` unless the time amount is specified in the service name.  It therefore *under*estimates the actual amount of services.


```{r}
waitlist_person <-
  waitlist_svs %>%
  group_by(ID) %>%
  summarize(
    first = min(first),
    last = max(last),
    Eligibility_Date = as.Date(max(Eligibility_Date)),
    svs_list = paste(SERVICE_DESC, collapse = ', '),
    hrs = sum(hrs)
  ) %>%
  mutate(
    # Most recent date in services data pulled
    max_date = max(svs$FROM_DATE),
    elig_interval = as.numeric(max_date - Eligibility_Date),
    svs_interval = as.numeric(last - first),
    pct_days_svd = round(svs_interval/elig_interval * 100, digits = 1)
  )

waitlist_person %>%
  select(ID,svs_list,elig_interval,svs_interval,pct_days_svd,hrs) %>%
  ungroup() %>%
  arrange(pct_days_svd) %>%
  datatable(
    rownames = FALSE,
    colnames = c('Enrollee ID','Services provided','Days on waitlist','Days receiving services','Percent days covered','Approximate hours'),
    caption = 'Service coverage for individuals on Autism Waiver waitlist, by Enrollee',
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  )

```

## Core Services per Month for Waitlisted Autism Enrollees

In order to ensure support for Autism Waiver enrollees who are currently on the waitlist for ABA services, we defined a complement of core services which should be provided each month to waitlisted individuals.  This set of services includes the receipt of one or more unit of each of the following services each month following an individual's eligibility date:

- Supports Coordination (`T1016`) or Targeted Case Management (`T1017`) 
- Community Living Supports (`H2015`) 
- Respite services (`T1005`)
- Family training (`0370T`)
- Medication review (`90792`,`99212:99215`)

```{r}

# Select field from WSA to join in later
wsa_elig_dt <- wsa %>% select(ID,Eligibility_Date)

min_waitlist_svs <-
  wsa %>%
  # Filter individuals who are on waitlist
  # Is this an accurate way to get at these individuals?
  filter(
    Status == "Open"
    & IPOSExists == F
    & is.na(Eligibility_Date) == F
  ) %>%
  select(ID,PIHP_CMH_Name,Referral_Date:Eligibility_End_Date) %>%
  left_join(svs, by = c("ID" = "MEDICAID_ID")) %>%
  filter(
    # Only include services occurring after eligibility for autism
    FROM_DATE >= as.Date(Eligibility_Date) 
    # or if there is no service date (to keep people who didn't receive svs)
    | is.na(FROM_DATE) == T 
  ) %>%
  filter(
    # Only include 'core' services defined by Region 10
    CPT_CD %in% c("T1016","T1017","H2015","T1005","0370T","90792",as.character(99201:99215))
  ) %>%
  mutate(
    # Create month dates for grouping
    month = floor_date(FROM_DATE, unit = "month")
  ) %>%
  droplevels() %>%
  ungroup() %>%
  group_by(month,ID,CPT_CD) %>%
  summarize(units = sum(UNITS, na.rm = T)) %>%
  ungroup() %>%
  # Create a dataframe with all combinations of the vars listed
  # in order to ensure that months without services are identified
  complete(month,ID,CPT_CD) %>%
  group_by(ID,month) %>%
  # Spread to show one row per person per month
  spread(CPT_CD,units) %>%
  # Join back to get eligibility date to filter months for each person that precede
  left_join(wsa_elig_dt, by = "ID") %>%
  filter(month >= floor_date(Eligibility_Date, unit = "month")) %>%
  # Convert NA values to zeroes
  mutate_at(
    .cols = vars(`0370T`:`T1017`), 
    .funs = funs(replace(., is.na(.), 0))
  ) %>%
  # Sum units for services to identify criteria
  mutate(
    cm = T1016 + T1017,
    cls = H2015,
    respite = T1005,
    med_rvw = `90792` + `99212` + `99213` + `99214` + `99215`,
    fam_svs = `0370T`
  ) %>%
  select(ID,month,cm,cls,respite,med_rvw,fam_svs) %>%
  mutate(
    min_svs = cm >= 1 & cls >= 1 & respite >= 1 & med_rvw >= 1 & fam_svs >= 1
  ) %>%
  arrange(ID, month)
```


# Conformance with Prescribed Hours for ABA Services

For individuals who are receiving autism waiver services, what is the amount of services and does it match their plan?  The bar chart below shows the percentage of authorized hours of ABA that have been provided for each active autism waiver enrollee over the past 90 days.  Individuals whose services fall outside of the acceptable range (*i.e. <75% or >125%*) of the prescribed hours are differentiated by color.

```{r}

#### Calculate ABA services per week per person ####

# Define ABA codes
aba_cpt <- c("S5108", # Old observation code
             "H2019", # Old treatment codes
             "0364T","0365T","0366T","0367T","0372T","0373T","0374T",
             "0368T","0369T") # Observation codes

aba_rx_hrs <- wsa_ipos %>% select(MEDICAID_ID = Beneficiary_ID,ABA_Hours)

ipos_start <- 
  wsa %>% 
  filter(Status == "Open" & Currently_Inactive == F) %>% 
  select(MEDICAID_ID = ID,IPOS_Start_Date)

last_90 <- as.Date(max(svs$FROM_DATE) - 90)

aba_svs <-
svs %>%
  filter(
    CPT_CD %in% aba_cpt
    & USED_MOD == "U5" # Only include U5 modifier for Autism services
  ) %>%
  droplevels() %>%
  mutate(
    # Recode unit type as numeric conversion factor
    hr_conv = recode(
      as.character(UNIT_TYPE),
      `15 Minutes` = '0.25', 
      `30 Minutes` = '0.5',
      `Hour` = '1', 
      `Up to 15 min` = '0.25',
      `Encounter` = 'NA'
    ),
    hr_conv = as.numeric(hr_conv),
    # Calculate hours
    hrs = UNITS * hr_conv,
    # Create week dates for grouping
    week = floor_date(FROM_DATE, unit = "week")
  ) %>%
  select(PROVIDER_NAME,MEDICAID_ID,CPT_CD:FROM_DATE,week,UNITS,hrs) 

aba_week <-
  aba_svs %>%
  # Get prescribed ABA hours from WSA IPOS
  left_join(aba_rx_hrs, by = "MEDICAID_ID") %>%
  # Get IPOS start date from WSA
  left_join(ipos_start, by = "MEDICAID_ID") %>%
  # Only include services if they occur after the IPOS Start Date
  filter(FROM_DATE >= IPOS_Start_Date) %>%
  # Only include services if they are in most recent 90 days of dataset
  filter(FROM_DATE >= last_90) %>%
  group_by(MEDICAID_ID,week) %>%
  summarize(
    hrs = sum(hrs, na.rm = T),
    ABA_Hours = max(ABA_Hours, na.rm = T)
  ) %>%
  mutate(
    # Calculate difference between actual and prescribed hours
    # Negative values indicate hours less than goal
    diff = hrs - ABA_Hours,
    # Calculate % of prescribed hours received
    pct = round(hrs / ABA_Hours * 100, digits = 1),
    # Tag Compliant and Non-Compliant weeks
    in_range_wk = ifelse(between(pct,75,125),T,F)
  ) %>%
  ungroup() %>%
  group_by(MEDICAID_ID) %>%
  mutate(
    pct_all = round(sum(hrs) / sum(ABA_Hours) * 100, digits = 1),
    in_range_all = ifelse(between(pct_all,75,125),T,F)
  ) 

aba_per <-
  aba_week %>%
  group_by(MEDICAID_ID) %>%
  summarize(
    hrs = sum(hrs, na.rm = T),
    rx_hrs = sum(ABA_Hours, na.rm = T)
  ) %>%
  mutate(
    pct = round(hrs / rx_hrs * 100, digits = 1),
    in_range = ifelse(between(pct,75,125),T,F)
  )

aba_per %>%
  mutate(
    MEDICAID_ID = fct_reorder(MEDICAID_ID,desc(pct)),
    in_range = ifelse(in_range == T,"Within range","Out of range")
  ) %>%
  plot_ly(x = ~MEDICAID_ID, y = ~pct, color = ~in_range, colors = c("#b0145b","#088da5")) %>%
  layout(
    title = "Percent of prescribed hours of ABA services received in last 90 days",
    xaxis = list(title = "Autism Waiver Enrollee",zeroline = F,showline = F,showticklabels = F,showgrid = F), 
    yaxis = list(title = "% of prescribed hours received"))


```

The individuals whose ABA service hours are outside of the prescribed range are shown in the table below:

```{r}
aba_per %>% 
  filter(in_range == F) %>% 
  select(-in_range) %>%
  arrange(desc(pct)) %>%
  datatable(
    rownames = FALSE,
    colnames = c('Enrollee ID','Hours of service','Prescribed Hours','Percent of prescribed hrs'),
    caption = 'Autism waiver enrollees receiving services outside of prescribed range'
  )
```


# Required Observation of ABA Services

According to the Autism Waiver requirements, for every 10 hours of treatment provided (*including codes from the following:* `0364T`:`0367T`,`0372T`:`0374T`,`H2019`) to an individual client, there should be 1 hour of observation (*including codes from the following:* `S5108`,`0368T`,`0369T`).  The table below shows the percentage of active autism waiver clients at each CMHSP who have received at least the required ratio of treatment to observation (10:1) during the most recent 90 days of available service data.  Individuals with fewer than 10 hours of treatment during this period are omitted from the calculations.

```{r}
# Note: uses filters defined above for service codes and clients included

observe <-
svs %>%
  filter(
    CPT_CD %in% aba_cpt
    & USED_MOD == "U5" # Only include U5 modifier for Autism services
  ) %>%
  droplevels() %>%
  mutate(
    # Recode unit type as numeric conversion factor
    hr_conv = recode(
      as.character(UNIT_TYPE),
      `15 Minutes` = '0.25', 
      `30 Minutes` = '0.5',
      `Hour` = '1', 
      `Up to 15 min` = '0.25',
      `Encounter` = 'NA'
    ),
    hr_conv = as.numeric(hr_conv),
    # Calculate hours
    hrs = UNITS * hr_conv,
    # Create week dates for grouping
    week = floor_date(FROM_DATE, unit = "week"),
    month = floor_date(FROM_DATE, unit = "month")
  ) %>%
  # Get prescribed ABA hours from WSA IPOS
  left_join(aba_rx_hrs, by = "MEDICAID_ID") %>%
  # Get IPOS start date from WSA
  left_join(ipos_start, by = "MEDICAID_ID") %>%
  # Only include services if they occur after the IPOS Start Date
  filter(week >= IPOS_Start_Date) %>%
  # Only include services if they are in most recent 90 days of dataset
  filter(FROM_DATE >= last_90) %>%
  mutate(
    type = recode(
      CPT_CD,
      `S5108` = "Observation",
      `0368T` = "Observation",
      `0369T` = "Observation",
      `0364T` = "Treatment",
      `0365T` = "Treatment",
      `0366T` = "Treatment",
      `0367T` = "Treatment",
      `0372T` = "Treatment",
      `0373T` = "Treatment",
      `0374T` = "Treatment",
      `H2019` = "Treatment"
    )
  ) %>%
  select(
    MEDICAID_ID,PROVIDER_NAME,type,CPT_CD,SERVICE_DESC,
    month,week,FROM_DATE,UNITS,hrs
  ) %>%
  group_by(MEDICAID_ID,PROVIDER_NAME,type) %>%
  summarize(hrs = sum(hrs)) %>%
  ungroup() %>%
  spread(type, hrs) %>%
  # Exclude cases with < 10 hours of treatment
  filter(Treatment >= 10) %>%
  mutate(
    obs_ratio = Observation / Treatment,
    meets_req = obs_ratio >= 0.1
  ) %>%
  group_by(PROVIDER_NAME) %>%
  summarize(
    meets_req = sum(meets_req, na.rm = T),
    n = n()
  ) %>%
  mutate(pct = round(meets_req / n * 100, digits = 1)) 

observe %>% 
  datatable(
    rownames = FALSE,
    colnames = c('CMHSP','Enrollees with required observation','All Enrollees','Percent meeting standard'),
    caption = 'Percentage of Autism waiver enrollees receiving required observation, by CMHSP',
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  ) %>%
  formatStyle(
    'pct',
    background = styleColorBar(observe$pct, 'grey'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )

```

