---
title: "Autism Waiver Compliance: Summary Report"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```

```{r}
# Load required packages
library(tidyverse); library(stringr); library(magrittr); library(lubridate); library(readxl)
library(DT); library(forcats); library(plotly); library(sparkline);library(feather)
```

```{r}

# Read in .csv files as dataframes
csv_path <- "data/"
svs <- read_feather(paste0(csv_path,"svs.feather"))
# Read in .csv files as dataframes
wsa <- read_feather(paste0(csv_path,"wsa.feather"))
# If re-opened, client is given new 'Case ID' but Medicaid ID (i.e. 'ID') should be same

# Read in data related to authorized units
wsa_ipos <- read_feather(paste0(csv_path,"wsa_ipos.feather"))

```

### This report was generated for Region 10 PIHP on `r max(svs$FROM_DATE)`

*Please note that the data in this report may not reflect the services of all enrollees that participated in the Autism Waiver Program.  Specifically, enrollee services may not show in the data for more recent periods where claims were not yet submitted.*

---

# Enrollees without a Behavioral Plan of Service

```{r}
wsa_tbl <-
wsa %>%
  filter(Status == "Open") %>%
  group_by(PIHP_CMH_Name) %>%
  summarize(
    without_IPOS = n() - sum(IPOSExists, na.rm = T),
    all = n()
  ) %>%
  mutate(
    pct = round(without_IPOS / all * 100, digits = 1)
  ) 

wsa_tbl %>%
  arrange(as.character(PIHP_CMH_Name)) %>%
  # Add subtotal line for PIHP
  rbind(
    ., 
    data.frame(
      PIHP_CMH_Name = "Region 10", 
      t(colSums(.[2:3])),
      pct = round(colSums(.[2])/colSums(.[3])*100,digits = 1)
    )
  ) %>%
  datatable(
    rownames = FALSE,
    colnames = c('CMHSP','Waitlisted Enrollees','Enrollees','Pct of Waitlisted Enrollees'),
    caption = 'Percentage of Autism waiver enrollees without IPOS, by CMHSP',
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  ) %>%
  formatStyle(
    'pct',
    background = styleColorBar(c(0,100), 'grey'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )
  
```

---

# Interim Services for Waitlisted enrollees

## Percent of Days Receiving Services per Waitlisted Enrollee

For each individual who is autism waiver eligible and on the waitlist, what percent of the days that the person has been on the waitlist are overlapped by the provision of non-ABA services (i.e. `Days receiving services`/`Days on waitlist`)?

```{r}
waitlist_svs <-
  wsa %>%
  # Filter individuals who are on waitlist
  filter(
    Status == "Open"
    & IPOSExists == F
    & is.na(Eligibility_Date) == F
  ) %>%
  select(ID,Case_ID,PIHP_CMH_Name,Referral_Date:Eligibility_End_Date) %>%
  left_join(svs, by = c("ID" = "MEDICAID_ID")) %>%
  filter(
    # Only include services occurring after eligibility for autism
    # Including a 90-day window for initial scheduling
    FROM_DATE > as.Date(Eligibility_Date) + 90    
    # or if there is no service date (to keep people who didn't receive svs)
    | is.na(FROM_DATE) == T 
  ) %>%
  group_by(Case_ID,PIHP_CMH_Name,CPT_CD,SERVICE_DESC,UNIT_TYPE,AUTISM_SRV,Eligibility_Date) %>%
  # Convert units to a common period
  mutate(
    # Recode unit type as numeric conversion factor
    hr_conv = recode(
      as.character(UNIT_TYPE),
      `15 Minutes` = '0.25', 
      `30 Minutes` = '0.5',
      `Hour` = '1', 
      `Up to 15 min` = '0.25',
      `Encounter` = 'NA'
    ),
    svs_conv = recode(
      as.character(SERVICE_DESC),
      `Psychotherapy, 60 min` = 1, 
      `Psychotherapy, 45 min` = 0.75,
      `Psychotherapy, 30 min` = 0.5
    ),
    hr_conv = ifelse(is.na(svs_conv) == F,svs_conv,as.numeric(hr_conv)),
    # Calculate hours
    hrs = UNITS * hr_conv
  ) %>%
  # What is the first and last date of each service?
  mutate(
    first = min(FROM_DATE),
    last = max(FROM_DATE)
  ) %>%
  # What is the number of units received?
  summarize(
    encounters = n(),
    units = sum(UNITS, na.rm = T),
    hrs = sum(hrs, na.rm = T),
    first = min(FROM_DATE),
    last = max(FROM_DATE)
  ) 

```


```{r}
waitlist_person <-
  waitlist_svs %>%
  group_by(Case_ID) %>%
  summarize(
    first = min(first),
    last = max(last),
    Eligibility_Date = as.Date(max(Eligibility_Date)),
    svs_list = paste(SERVICE_DESC, collapse = ', '),
    encounters = sum(encounters),
    autism_waiver_svs = sum(AUTISM_SRV, na.rm = T)
  ) %>%
  mutate(
    # Most recent date in services data pulled
    max_date = max(svs$FROM_DATE),
    elig_interval = as.numeric(max_date - (Eligibility_Date + 90)),
    svs_interval = as.numeric(last - first),
    pct_days_svd = round(svs_interval/elig_interval * 100, digits = 1),
    autism_waiver_svs = autism_waiver_svs > 0
  )

waitlist_person %>%
  plot_ly(
    x = ~pct_days_svd,
    hoverinfo = 'y+text',
    text = ~paste0("# of waitlisted individuals received services<br>",
                   "around this much of the time they were on the waitlist")
  ) %>%
  layout(
    xaxis = list(title = "% of eligible days receiving services"), 
    yaxis = list(title = "# of waitlisted individuals"), 
    showlegend = FALSE
  )

```

# Conformance with Prescribed Hours for ABA Services

What percent of individuals have received ABA services which fall outside of the acceptable range (*i.e. <75% or >125%*) of the prescribed hours?

```{r}

#### Calculate ABA services per week per person ####

# Define ABA codes
aba_cpt <- c("S5108", # Old observation code
             "H2019", # Old treatment codes
             "0364T","0365T","0366T","0367T","0372T","0373T","0374T",
             "0368T","0369T") # Observation codes

aba_rx_hrs <- wsa_ipos %>% select(MEDICAID_ID = Beneficiary_ID,Case_ID,ABA_Hours)

ipos_start <- 
  wsa %>% 
  filter(Status == "Open" & Currently_Inactive == F) %>% 
  select(MEDICAID_ID = ID,IPOS_Start_Date)

last_90 <- as.Date(max(svs$FROM_DATE) - 90)

aba_svs <-
svs %>%
  filter(
    CPT_CD %in% aba_cpt
    & USED_MOD == "U5" # Only include U5 modifier for Autism services
  ) %>%
  # Now remove observation codes
  filter(!CPT_CD %in% c("S5108","0368T","0369T")) %>%
  droplevels() %>%
  mutate(
    # Recode unit type as numeric conversion factor
    hr_conv = recode(
      as.character(UNIT_TYPE),
      `15 Minutes` = '0.25', 
      `30 Minutes` = '0.5',
      `Hour` = '1', 
      `Up to 15 min` = '0.25',
      `Encounter` = 'NA'
    ),
    hr_conv = as.numeric(hr_conv),
    # Calculate hours
    hrs = UNITS * hr_conv,
    # Create week dates for grouping
    week = floor_date(FROM_DATE, unit = "week")
  ) %>%
  select(PROVIDER_NAME,MEDICAID_ID,CPT_CD:FROM_DATE,week,UNITS,hrs) 

aba_summary <-
  aba_svs %>%
  # Get prescribed ABA hours from WSA IPOS
  left_join(aba_rx_hrs, by = "MEDICAID_ID") %>%
  # Get IPOS start date from WSA
  left_join(ipos_start, by = "MEDICAID_ID") %>%
  # Only include services if they occur after the IPOS Start Date
  filter(FROM_DATE >= IPOS_Start_Date) %>%
  # Only include services if they are in most recent 90 days of dataset
  filter(FROM_DATE >= last_90) %>%
  group_by(Case_ID,PROVIDER_NAME,week) %>%
  summarize(
    hrs = sum(hrs, na.rm = T),
    ABA_Hours = max(ABA_Hours)
  ) %>%
  mutate(
    # Calculate difference between actual and prescribed hours
    # Negative values indicate hours less than goal
    diff = hrs - ABA_Hours,
    # Calculate % of prescribed hours received
    # Tag Compliant and Non-Compliant weeks
   
    pct_all = round(sum(hrs) / sum(ABA_Hours) * 100, digits = 1),
    in_range_all = ifelse(between(pct_all,75,125),T,F))%>%
  ungroup() %>%
  group_by(Case_ID,PROVIDER_NAME) %>%
  summarise(
    hrs = sum(hrs, na.rm = T),
    ABA_Hours = sum(ABA_Hours, na.rm = T)
  )%>%
  mutate(
    pct = round(hrs / ABA_Hours * 100, digits = 1),
    in_range = ifelse(between(pct,75,125),T,F))%>%
  group_by(PROVIDER_NAME)%>%
  summarise(
    numerator=sum(in_range),
    denominator=n()
  ) %>%
  mutate(
    pct=round(numerator/denominator*100,digits=1)
  )

aba_summary %>% 
  arrange(as.character(PROVIDER_NAME)) %>%
  # Add subtotal line for PIHP
  rbind(
    ., 
    data.frame(
      PROVIDER_NAME = "Region 10", 
      t(colSums(.[2:3])),
      pct = round(colSums(.[2])/colSums(.[3])*100,digits = 1)
    )
  ) %>%
  datatable(
    rownames = FALSE,
    colnames = c('CMHSP','Enrollees within prescribed range','Enrollees receiving ABA Services','Percent meeting standard'),
    caption = 'Percentage of Autism waiver enrollees receiving prescribed ABA hours, by CMHSP for the last 90 days',
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  ) %>%
  formatStyle(
    'pct',
    backgroundSize = '100% 90%',
    background = styleColorBar(c(0,100), 'grey'),
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )

```

# Required Observation of ABA Services



```{r}
# Note: uses filters defined above for service codes and clients included

observe <-
svs %>%
  filter(
    CPT_CD %in% aba_cpt
    & USED_MOD == "U5" # Only include U5 modifier for Autism services
  ) %>%
  droplevels() %>%
  mutate(
    # Recode unit type as numeric conversion factor
    hr_conv = recode(
      as.character(UNIT_TYPE),
      `15 Minutes` = '0.25', 
      `30 Minutes` = '0.5',
      `Hour` = '1', 
      `Up to 15 min` = '0.25',
      `Encounter` = 'NA'
    ),
    hr_conv = as.numeric(hr_conv),
    # Calculate hours
    hrs = UNITS * hr_conv,
    # Create week dates for grouping
    week = floor_date(FROM_DATE, unit = "week"),
    month = floor_date(FROM_DATE, unit = "month")
  ) %>%
  # Get prescribed ABA hours from WSA IPOS
  left_join(aba_rx_hrs, by = "MEDICAID_ID") %>%
  # Get IPOS start date from WSA
  left_join(ipos_start, by = "MEDICAID_ID") %>%
  # Only include services if they occur after the IPOS Start Date
  filter(week >= IPOS_Start_Date) %>%
  # Only include services if they are in most recent 90 days of dataset
  filter(FROM_DATE >= last_90) %>%
  mutate(
    type = recode(
      CPT_CD,
      `S5108` = "Observation",
      `0368T` = "Observation",
      `0369T` = "Observation",
      `0364T` = "Treatment",
      `0365T` = "Treatment",
      `0366T` = "Treatment",
      `0367T` = "Treatment",
      `0372T` = "Treatment",
      `0373T` = "Treatment",
      `0374T` = "Treatment",
      `H2019` = "Treatment"
    )
  ) %>%
  select(
    Case_ID,PROVIDER_NAME,type,CPT_CD,SERVICE_DESC,
    month,week,FROM_DATE,UNITS,hrs
  ) %>%
  group_by(Case_ID,PROVIDER_NAME,type) %>%
  summarize(hrs = sum(hrs)) %>%
  ungroup() %>%
  mutate(hrs = ifelse(is.na(hrs),0,hrs)) %>%
  spread(type, hrs) %>%
  mutate(
    obs_ratio = round(Observation / Treatment, digits = 2),
    meets_req = obs_ratio >= 0.1 & obs_ratio <= 1.0
  ) %>%
  # Exclude cases with < 10 hours of treatment if they don't meet criteria
  filter(Treatment >= 10 | (Treatment < 10 & meets_req == T))

observe_agg <-
observe %>%
  group_by(PROVIDER_NAME) %>%
  summarize(
    meets_req = sum(meets_req, na.rm = T),
    n = n()
  ) %>%
  mutate(pct = round(meets_req / n * 100, digits = 1)) 
  
observe_agg %>% 
  arrange(as.character(PROVIDER_NAME)) %>%
  # Add subtotal line for PIHP
  rbind(
    ., 
    data.frame(
      PROVIDER_NAME = "Region 10", 
      t(colSums(.[2:3])),
      pct = round(colSums(.[2])/colSums(.[3])*100,digits = 1)
    )
  ) %>%
  datatable(
    rownames = FALSE,
    colnames = c('CMHSP','Enrollees with required observation','Enrollees receiving ABA Services','Percent meeting standard'),
    caption = 'Percentage of Autism waiver enrollees receiving required observation, by CMHSP',
    extensions = c('Responsive','Buttons'),
    options = list(
      dom = 't',
      buttons = c('colvis')
    )
  ) %>%
  formatStyle(
    'pct',
    background = styleColorBar(c(0,100), 'grey'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )

```

